Legion: Expressing Locality and Independence with Logical Regions

### 2012-05-04 doc/sc2012/mapping.tex
mapping

\section{Mapping}
\label{sec:mapping}

While the Legion programming model describes locality and independence information abstractly,
to run a Legion application on a machine we must make concrete decisions about
where tasks will be run and where region instances will be placed.  Instead of placing
the burden of decision making on the programming system, we introduce a programmable mapping
interface that gives a programmer total control over how these decisions are made.  Armed
with this interface, the programmer can then specify application- or machine-specific
mapping decisions that would be difficult for a general-purpose programming system to infer.
In this section we describe the general mapping interface (Section \ref{sec:mapinterface}),
our base implmentation of this interface (Section \ref{sec:defmapper}), and the benefits
of creating custom mappers (Section \ref{sec:custommap}).

\subsection{Mapping Interface}
\label{sec:mapinterface}
A mapper is any object that implements the mapping interface.  The mapping interface
consists of ten function calls that the runtime will make to request direction from
the mapper about how a program should be executed.  To aid in performing the 
decisions made in these calls the mapper has access to a simple 
interface for inspecting the properties of the machine on which it is running.  This includes 
information about all the processors and their kinds (i.e. CPU,GPU), as well as the memories visible
to each processor and their latencies and bandwidths.  For brevity we only
discuss three of the mapping interface calls.

\begin{itemize}
\item {\tt select\_initial\_processor} - for each task the runtime system will
ask the mapper to which processor the task should be sent.  The mapper 
can choose to keep the task on the local processor or send it to any other processor
in the system.

\item {\tt permit\_task\_steal} - when handling a steal request from another processor
the runtime will ask the mapper to decide which tasks are allowed to be stolen.  
By always returning an empty set, the mapper can disable stealing.

\item {\tt map\_task\_region} - for each logical region requested for a task, the
runtime will query the mapper for a prioritized list of memories where a physical instance for this
task should be placed.  To aid the mapping decision the runtime provides a list of the current valid physical intances
of the logical region.  The mapper returns a list of memories in which the runtime
should attempt to either reuse or manifest a physical instance of the logical region.  The
runtime traverses this list and searches for a currently valid instance.  If none can be
found, it attempts to allocate a physical instance and then issue the necessary copies
to retrive the valid data.  If both options fail, it moves on to the next memory in the list.
\end{itemize}

There are two very elegant properties of this mapping interface.  The first property is
that all decisions made by a mapper are orthogonal to the correctness of the program and can only 
impact performance.  Regardless of where a mapper chooses to place a task or map a region, 
the runtime will always schedule tasks and copies in accordance with the privileges and 
coherence properties specified in the Legion program.  Therefore when writing an application
in Legion, a programmer can begin by using the default mapper and later optimize his program
by creating a custom mapper and gradually refining its mapping decisions.

The second useful property of the mapping interface is that it isolates machine-specific decisions
to a specific module of code.  As a result, Legion programs are highly
portable.  To port a Legion program to a new architecture, a programmer only has to
implement a new mapper with decisions specific to the new architecture.

\subsection{Default Mapper}
\label{sec:defmapper}
To make writing Legion applications easier, we provide a default mapper implementation that
can be used to get an application working quickly at a moderate performance level.  The 
default mapper employs a simple scheme for mapping tasks and regions.  When
{\tt select\_initial\_processor} is invoked for the default mapper, it first checks for which kind of processors
the task has variants (i.e. GPU).  If the fastest variant is for the kind of processor the mapper is managing
it will choose to keep the task local, otherwise it will send the task to the closest processor of the
fastest variant kind.

The default mapper employs a Cilk-like algorithm for task stealing\cite{CILK95}.  Tasks are kept local
to their target processor whenever possible and only moved when stolen.  Unlike Cilk, the default mapper
has the information necessary for locality-aware stealing.  When {\tt permit\_task\_steal} is called for a task, 
the default mapper inspects the logical regions for the task being stolen, and will mark that other tasks using 
the same logical regions should be stolen as well.

For calls to {\tt map\_task\_region}, the default mapper constructs a stack of memories ordered from best-to-worst
by bandwidth from the local processor.  This stack is then returned as the location of memories to be used for
mapping each region.  Note that this is a greedy algorithm that works well for the common case, but can cause some 
regions to be pulled unnecessarily close to the processor and consume precious fast memory space.

@@

Each custom mapper extends the default mapper.  A programmer only has to override the functions necessary 
for creating a custom mapping.  Custom mappers can be used to create totally static mappings, 
mappings that memoize their results, or even totally dynamic mappings.  We describe examples of 
custom mappers in Section \ref{sec:exp}.

>>>>

\section{Mapping Interface}
\label{sec:mapping}

As mentioned previously, a novel aspect of Legion is the {\em mapping interface},
which gives programmers control over where tasks run and where region instances are placed,
making possible application- or machine-specific mapping decisions that would be difficult for a 
general-purpose programming system to infer.
We describe the interface (Section \ref{sec:mapinterface}),
our base implementation (Section \ref{sec:defmapper}), and the benefits
of creating custom mappers (Section \ref{sec:custommap}).

\subsection{The Interface}
\label{sec:mapinterface}

The mapping interface consists of ten methods that SOOPs call for
mapping decisions.  A {\em mapper} implementing these methods has access to a simple interface
for inspecting properties of the machine, including a list of
processors and their type (i.e. CPU, GPU) and a list of
memories visible to each processor and their latencies and bandwidths.
For brevity we only discuss three of the mapping interface calls:

\begin{itemize}
\item {\tt select\_initial\_processor} - For each task $t$ in its mapping queue a SOOP will
ask for a processor for $t$.  The mapper can keep the task on the local processor or send it to 
any other processor in the system.

\item {\tt permit\_task\_steal} - When handling a steal request a SOOP
asks which tasks may be stolen.  Stealing can be disabled by always returning the empty set.

\item {\tt map\_task\_region} - For each logical region $r$ used by a task, a
SOOP asks for a prioritized list of memories where a physical instance of $r$ should be placed.  
The SOOP provides a list of $r$'s current valid physical instances;
the mapper returns a priority list of memories in which the SOOP
should attempt to either reuse or create a physical instance of $r$.  Beginning with
the first memory, the SOOP uses a current valid instance if one is present.  Otherwise,
the SOOP attempts to allocate a physical instance and issue copies
to retrieve the valid data.  If that also fails, the SOOP moves on to the next memory in the list.
\end{itemize}

The mapping interface has two desirable properties.  First, all mapper
decisions are orthogonal to program correctness and can only impact
performance.  Regardless of where a mapper places a task or region,
the SOOPs schedule tasks and copies in accordance with the privileges
and coherence properties specified in the program.  Therefore when
writing a Legion application, a programmer can begin by using the
default mapper and later improve performance by creating and refining a custom mapper.
Second, the mapping interface isolates machine-specific decisions
to the mapper.  As a result, Legion programs are highly
portable.  To port a Legion program to a new architecture, a programmer need only
implement a new mapper with decisions specific to the new architecture.

\subsection{Default Mapper}
\label{sec:defmapper}
To make writing Legion applications easier, we provide a default
mapper that can quickly get an application working
with moderate performance.  The default mapper employs a
simple scheme for mapping tasks.  When {\tt
select\_initial\_processor} is invoked, the mapper
checks the type of processors for which task $t$ has
implementations (e.g., GPU).  If the fastest implementation is for the
local processor the mapper keeps $t$ local, otherwise
it sends $t$ to the closest processor of the fastest kind that can run
$t$.

The default mapper employs a Cilk-like algorithm for task
stealing \cite{CILK95}.  Tasks are kept local 
whenever possible and only moved when stolen.  Unlike Cilk, the
default mapper has the information necessary for locality-aware
stealing.  When {\tt permit\_task\_steal} is called for a task, the
default mapper inspects the logical regions for the task being stolen
and marks that other tasks using the same logical regions should
be stolen as well.

For calls to {\tt map\_task\_region}, the default mapper constructs a stack of memories ordered from best-to-worst
by bandwidth from the local processor.  This stack is then returned as the location of memories to be used for
mapping each region.  This greedy algorithm works well in common cases, but can cause some 
regions to be pulled unnecessarily close to the processor, consuming precious fast memory.

@@

Each custom mapper extends the default mapper.  A programmer need only override the functions necessary 
for creating a custom mapping.  Custom mappers can be used to create totally static mappings, 
mappings that memoize their results, or even totally dynamic mappings.  We describe examples of 
custom mappers in Section \ref{sec:exp}.

### 2012-05-04 doc/sc2012/soop.tex
mapping

Figure~\ref{fig:gpumapping} shows a timeline of our mapping of the
circuit simulation on a cluster of GPUs.  The $i$th private node and
wire regions are mapped to the $i$th GPU's DRAM and simply left there
for all three phases of the simulation.  The $i$th shared and ghost
regions are mapped to zero-copy memory most of the time, but, as just
illustrated, must be reduced out to the all-shared node region in the
global memory (which is managed by GASNet) and then a new instance of the shared
subregions created with the updated information before the {\tt
  update\_voltages} tasks execute.  A similar situation arises before
{\tt calc\_new\_currents} tasks run again: the updates to the shared
nodes by {\tt update\_voltage} must be propagated to the ghost nodes
before continuing.

\begin{figure}
\includegraphics[scale=0.48]{figs/CircuitMem.pdf}
\caption{Tasks and data for the circuit simulation on a cluster of GPUs.}
\label{fig:gpumapping}
\end{figure}

>>>>

%
% FIXME   CUT
%
%Figure~\ref{fig:gpumapping} shows a timeline of our mapping of the
%circuit simulation on a cluster of GPUs.  The $i$th private node and
%wire regions are mapped to the $i$th GPU's DRAM and simply left there
%for all three phases of the simulation.  The $i$th shared and ghost
%regions are mapped to zero-copy memory most of the time, but, as just
%illustrated, must be reduced out to the all-shared node region in the
%global memory (which is managed by GASNet) and then a new instance of the shared
%subregions created with the updated information before the {\tt
%  update\_voltages} tasks execute.  A similar situation arises before
%{\tt calc\_new\_currents} tasks run again: the updates to the shared
%nodes by {\tt update\_voltage} must be propagated to the ghost nodes
%before continuing.
%
%\begin{figure}
%\includegraphics[scale=0.48]{figs/CircuitMem.pdf}
%\caption{Tasks and data for the circuit simulation on a cluster of GPUs.}
%\label{fig:gpumapping}
%\end{figure}
